DATA ANALYSIS HW4

PROBLEM 1

```{r}
library(DAAG)
data("ais")

# A: Linear model with lbm as the response and sport as the predictor
model_lbm_sport <- lm(LBM ~ Sport, data = ais)
summary(model_lbm_sport)

# B: Boxplots for lbm according to sport
boxplot(LBM ~ Sport, data = ais, 
        main = "Boxplot of LBM by Sport",
        xlab = "Sport",
        ylab = "Lean Body Mass")
```
C: Interpretation of the F-test -
The F-test in the summary output is used to determine whether there is a 
significant effect of sport on lean body mass across the different sports. 
The null hypothesis for the F-test is that all group means are equal, or in 
other words, sport has no effect on lean body mass.

The provided output shows an F-statistic value of 12.19 with a p-value of 
3.562e-15. The very small p-value (much less than any common significance level 
like 0.05 or 0.01) leads us to reject the null hypothesis. This means there is 
strong evidence that there are differences in lean body mass among at least some
of the sports categories.

D: Interpretation of the coefficient estimate for Field and Tennis -
The coefficient estimate for each sport represents the difference in lean body 
mass between athletes of that sport and the reference category (which is not 
shown and is likely the one not listed, often the one alphabetically first if 
the reference was not manually set).

For sportField, the estimate is 8.2014 with a p-value of 0.01234. Since the 
p-value is less than 0.05, it suggests that athletes in track and field sports 
have, on average, 8.2014 units higher lean body mass compared to the reference 
sport category, holding all else equal, and this difference is statistically 
significant.

For Tennis, the estimate is -12.0173 with a p-value of 0.00213. This negative 
coefficient indicates that tennis athletes have, on average, 12.0173 units lower
lean body mass compared to the reference sport category, holding all else equal,
and this difference is statistically significant.

To sum up, the lean body mass for track and field athletes is significantly 
higher than the reference group, while for tennis athletes, it's significantly 
lower.

```{r}
# E: Estimate the contrast between Tennis and W_Polo
library(emmeans)

# Estimate the emmeans for 'sport'
emm <- emmeans(model_lbm_sport, specs = "Sport")

# Calculate the pairwise contrasts
contrast_estimate <- pairs(emm)

# We can use the summary function to extract the contrasts
summary_contrast <- summary(contrast_estimate)

# Find the row in the summary that corresponds to the Tennis vs. W_Polo contrast
tennis_vs_wpolo_row <- which(summary_contrast$contrast == "tennis - w_polo")

# Obtain the 90% CI for the Tennis vs. W_Polo contrast
tennis_vs_wpolo_ci <- confint(contrast_estimate, level = 
                                0.90)[tennis_vs_wpolo_row, ]

# Output the results
print(tennis_vs_wpolo_ci)
 
# F: Adding the ht (height) variable to the model
# Fit the model including height
model_lbm_sport_ht <- lm(LBM ~ Sport + Ht, data = ais)

# Get a summary of the new model to interpret the coefficients
summary(model_lbm_sport_ht)

```
Interpretation of the coefficient for ht:
The coefficient estimate for ht is 1.12073, with a highly significant p-value 
(p < 2.2e-16). This suggests a strong relationship between height and lean body 
mass. The interpretation is that for each one-centimeter increase in height, the
lean body mass is expected to increase by an average of 1.12073 kilograms, 
assuming all other factors remain constant.

Interpretation of the coefficient for Field now:
The coefficient for Field (assuming it represents Track and Field athletes) is 
17.23481 with a p-value < 2.2e-16, which is also highly significant. This 
suggests that, holding height constant, Track and Field athletes have a lean 
body mass that is, on average, 17.23481 kilograms more than the reference 
category.

PROBLEM 2

```{r}
# A: Fit the model with fertility as the response variable
library(alr4)
data("UN11")

model_fertility <- lm(fertility ~ ppgdp + lifeExpF + pctUrban, data = UN11)
summary(model_fertility)

# B: Plot of fitted values vs. residuals
plot(fitted(model_fertility), residuals(model_fertility), 
     main="Fitted Vs Residual Values",
     xlab="Fitted Values",
     ylab="Residuals")
abline(h = 0, col = "red")
```
Interpretation: The residual plot of the linear model shows no clear patterns, 
which implies the linear model might be a suitable fit, as it doesn't seem to 
violate the linearity assumption significantly.

```{r}
# C: Scatter plot pairs
pairs(~fertility + ppgdp + lifeExpF + pctUrban, data = UN11)
```
In the scatterplot examining fertility versus per capita GDP, a negative trend 
emerges, indicating that fertility rates tend to fall as per capita GDP climbs. 
This trend appears to curve, suggesting a nonlinear relationship. 

When observing fertility against female life expectancy, a negative link is also
present; nations with longer female life expectancies often have lower fertility
rates. The pattern here seems fairly straight, hinting at a linear association.

The connection between fertility and the percentage of the urban population 
isn’t completely linear. Initially, there's considerable variability, yet a
general negative direction is noticeable, with fertility rates tending to drop 
as the urban population proportion increases.

Looking at per capita GDP versus female life expectancy, a positive association 
is discernible. Higher per capita GDP often aligns with longer female life 
expectancy, and the pattern of this relationship seems to curve, implying 
nonlinearity.

The relationship between per capita GDP and the percentage of the urban
population is also positive; higher GDP per capita is linked with a larger urban
population share, with the relationship again being nonlinear.

Lastly, female life expectancy correlates positively with the percentage of 
urban population; countries with more urbanized populations generally see higher
female life expectancies.

The observations of nonlinear trends suggest that simple linear models might not
adequately capture the nuances of these relationships without transformations of
the variables, such as applying a logarithmic scale to per capita GDP. 
Additionally, the presence of positive correlations among per capita GDP, female
life expectancy, and the urban population percentage raises the issue of 
potential multicollinearity, which could complicate the interpretation of these 
variables in a multivariate model.
```{r}
# D: Model with log(ppgdp) and related plots
model_fertility_log_ppgdp <- lm(fertility ~ log(ppgdp) + lifeExpF + pctUrban, 
                                data = UN11)
summary(model_fertility_log_ppgdp)

pairs(~fertility + log(ppgdp) + lifeExpF + pctUrban, data = UN11)

plot(fitted(model_fertility_log_ppgdp), residuals(model_fertility_log_ppgdp), 
     xlab = "Fitted Values", ylab = "Residuals", 
     main = "Residuals vs Fitted Values with log(ppgdp)")
abline(h = 0, col = "red")
```
The regression analysis shows a significant inverse correlation between a 
country's GDP per capita and its fertility rate. A one-unit increase in the 
logged GDP per capita correlates with a 0.1979 unit decrease in the fertility 
rate, assuming female life expectancy and urbanization rates remain constant. 
The improved linearity in scatter plots after logging GDP suggests that the 
transformed model captures the relationship between the variables more 
effectively. While the spread of residuals still exhibits some variability, the 
previous cone-shaped pattern is less pronounced, marking an advancement over the
initial model. However, residual analysis indicates room for further refinement.
```{r}
# E:  Model with both log(fertility) and log(ppgdp)
model_log_fertility_log_ppgdp <- lm(log(fertility) ~ log(ppgdp) + lifeExpF + 
                                      pctUrban, data = UN11)
summary(model_log_fertility_log_ppgdp)

pairs(~log(fertility) + log(ppgdp) + lifeExpF + pctUrban, data = UN11)

plot(fitted(model_log_fertility_log_ppgdp), 
     residuals(model_log_fertility_log_ppgdp), 
     xlab = "Fitted Values", ylab = "Residuals", 
     main = "Residuals vs Fitted Values with log(fertility) and log(ppgdp)")
abline(h = 0, col = "red")
```
The regression analysis indicates a significant negative association between a 
country's GDP per capita and its fertility rate; a one-unit increase in the 
log-transformed GDP per capita is associated with a 0.0758 unit decrease in 
fertility rate, holding female life expectancy and urbanization constant. The 
residual plot shows a marked improvement with a more uniform variance, 
suggesting an appropriate model fit without discernible trends.

PROBLEM 4

```{r}
# Part A:
# Load the Swiss data
data(swiss)

# Full model with all predictors
full_model <- lm(Fertility ~ ., data = swiss)

# Forward selection
forward_model <- step(object = lm(Fertility ~ 1, data = swiss), 
                      scope = list(lower = formula(lm(Fertility ~ 1, 
                                                      data = swiss)), 
                                   upper = formula(full_model)), 
                      direction = "forward", 
                      trace = 0)

# Backward selection
backward_model <- step(full_model, direction = "backward", trace = 0)

# Step wise selection (both forward and backward steps)
stepwise_model <- step(object = lm(Fertility ~ 1, data = swiss), 
                       scope = list(lower = formula(lm(Fertility ~ 1, 
                                                       data = swiss)), 
                                    upper = formula(full_model)), 
                       direction = "both", 
                       trace = 0)

forward_model
backward_model
stepwise_model

```

Part B(i):
The models converge on Education, Catholic, and Infant.Mortality as key 
predictors for Fertility, though Agriculture's inclusion in only two models 
hints at its less definitive predictive role.

Part B(ii):
Education, Catholicism, and Infant Mortality emerge as the most consistent 
predictors of Fertility across all models, highlighting their strong correlation
with Fertility in Swiss provinces. Education inversely correlates with Fertility, suggesting that higher educational attainment may lead to lower Fertility rates.
A higher percentage of Catholics correlates with increased Fertility, while 
greater Infant Mortality rates also correspond with higher Fertility, 
potentially indicating compensatory reproductive behavior. These persistent 
associations across different models reinforce the significance of these factors
in understanding Fertility trends.
